---
title: "Homework 3 Question 1"
output: html_document
date: '2022-06-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dlm)
```

### Part A

```{r fig.width=10}
F_t <- 1.2
G_t <- 0.8
V_t <- 9
W_t <- 4
m_0 <- 0
C_0 <- 25

x <- read_csv("DLM_Data.csv", show_col_types = F)

filtered_data <- dlmFilter(
  y = x$yt,
  mod = dlm(
    FF = F_t,
    GG = G_t,
    V = V_t,
    W = W_t,
    m0 = m_0,
    C0 = C_0
  ),
  simplify = T
)

x <- x %>%
  mutate(
    a = filtered_data$a,
    R = unlist(dlmSvd2var(
      filtered_data$U.R,
      filtered_data$D.R
    ))
  )

x$a[40]
x$R[40]

x %>%
  ggplot(aes(time)) +
  geom_line(aes(y = yt, color = 'red')) +
  geom_line(aes(y = a, color = 'blue')) +
  geom_ribbon(
    aes(
      ymin = a - 1.96 * sqrt(R),
      ymax = a + 1.96 * sqrt(R)
    ),
    alpha = 0.2
  ) +
  scale_color_manual(
    labels = c("Predicted", "Observed"),
    values = c("red", "blue")
  ) +
  labs(
    title = expression(paste(
      "One-Step-Ahead Predictions of ",
      theta[t],
      " w/ Standard Errors"
    )),
    y = "Observations and Predictions",
    x = "Time",
    color = NULL
  )
```

We report the following numerical values:

<center>$a_{40}=3.53$</center>
<center>$R_{40}=5.95$</center>

### Part B

```{r fig.width=10}
x <- x %>%
  mutate(
    f = filtered_data$f,
    Q = F_t*R*F_t + V_t
  )

x$f[40]
x$Q[40]

x %>%
  ggplot(aes(time)) +
  geom_line(aes(y = yt, color = 'red')) +
  geom_line(aes(y = f, color = 'blue')) +
  geom_ribbon(
    aes(
      ymin = f - 1.96 * sqrt(Q),
      ymax = f + 1.96 * sqrt(Q)
    ),
    alpha = 0.2
  ) +
  scale_color_manual(
    labels = c("Predicted", "Observed"),
    values = c("red", "blue")
  ) +
  labs(
    title = expression(paste(
      "One-Step-Ahead Predictions of ",
      y[t],
      " w/ Standard Errors"
    )),
    y = "Observations and Predictions",
    x = "Time",
    color = NULL
  )

```

We report the following numerical values:

<center>$f_{40}=4.23$</center>
<center>$Q_{40}=17.57$</center>

### Part C

```{r fig.width=10}
x <- x %>%
  mutate(
    m = dropFirst(filtered_data$m),
    C = dropFirst(unlist(dlmSvd2var(
        filtered_data$U.C,
        filtered_data$D.C
      )))
  )

x$m[40]
x$C[40]

x %>%
  ggplot(aes(time)) +
  geom_line(aes(y = yt, color = 'red')) +
  geom_line(aes(y = m, color = 'blue')) +
  geom_ribbon(
    aes(
      ymin = m - 1.96 * sqrt(C),
      ymax = m + 1.96 * sqrt(C)
    ),
    alpha = 0.2
  ) +
  scale_color_manual(
    labels = c("Filtered", "Observed"),
    values = c("red", "blue")
  ) +
  labs(
    title = expression(paste(
      "Mean of the filtering distribution of ",
      theta[t],
      " w/ Standard Errors"
    )),
    y = "Observations and Filtering",
    x = "Time",
    color = NULL
  )
```

We report the following numerical values:

<center>$m_{40}=0.50$</center>
<center>$C_{40}=3.05$</center>

### Part D

$F_t=1.2$\
$G_t=0.8$\
$\sigma_v^2=9$\
$\sigma_{\omega}^2=4$\

$m_{22}=3.539$\
$c_{22}=3.048$\

$a_{23}=0.8(3.539)=2.831$\
$a_{24}=0.8(2.831)=2.265$\
$a_{25}=0.8(2.265)=1.812$\
$a_{26}=0.8(1.812)=1.450$\
$a_{27}=0.8(1.450)=1.160$\
$a_{28}=0.8(1.160)=0.928$\

$R_{23}=0.8(3.048)0.8+4=5.951$\
$Q_{23}=1.2(5.951)1.2+9=17.570$\
$C_{23}=(5.951)-(5.951)1.2(17.570)1.2(5.951)=-890.062$\

$R_{24}=0.8()0.8+4=$\
$Q_{24}=1.2()1.2+9=$\
$C_{24}=()-()1.2()1.2()=$\
$R_{25}=0.8()0.8+4=$\
$Q_{25}=1.2()1.2+9=$\
$C_{25}=()-()1.2()1.2()=$\
$R_{26}=0.8()0.8+4=$\
$Q_{26}=1.2()1.2+9=$\
$C_{26}=()-()1.2()1.2()=$\
$R_{27}=0.8()0.8+4=$\
$Q_{27}=1.2()1.2+9=$\
$C_{27}=()-()1.2()1.2()=$\
$R_{28}=0.8()0.8+4=$\
$Q_{28}=1.2()1.2+9=$\
$C_{28}=()-()1.2()1.2()=$\

### Part E

```{r fig.width=10}
smoothed_data <- dlmSmooth(filtered_data)

x <- x %>%
  mutate(
    theta = dropFirst(smoothed_data$s),
    S = dropFirst(unlist(dlmSvd2var(
      smoothed_data$U.S,
      smoothed_data$D.S
    )))
  )

x %>%
  ggplot(aes(time)) +
  geom_line(aes(y = yt, color = 'red')) +
  geom_line(aes(y = theta, color = 'blue')) +
  geom_ribbon(
    aes(
      ymin = theta - 1.96 * sqrt(S),
      ymax = theta + 1.96 * sqrt(S)
    ),
    alpha = 0.2
  ) +
  scale_color_manual(
    labels = c("Smoothed", "Observed"),
    values = c("red", "blue")
  ) +
  labs(
    title = expression(paste(
      "Smoothed Values of ",
      theta[t],
      " w/ Standard Errors"
    )),
    y = "Observations and Smoothing",
    x = "Time",
    color = NULL
  )
```

We report the following $\theta_t$ values for all $t$ such that $y_t$ is missing:

```{r echo=FALSE, results='asis'}
x %>%
  filter(is.na(yt)) %>%
  select(time, theta) %>%
  knitr::kable(align = 'll')
```

### Part F

```{r fig.width=10}
forecast <- dlmForecast(filtered_data, 10)

for (i in 1:10) {
  x[nrow(x) + 1,] = rep(NA, ncol(x))
  x$time[i + 100] = i + 100
}

x$a[101:110] = forecast$a[,1]
x$R[101:110] = unlist(forecast$R)
x$f[101:110] = forecast$f[,1]
x$Q[101:110] = unlist(forecast$Q)

x %>%
  ggplot(aes(time)) +
  geom_line(aes(y = yt, color = 'red')) +
  geom_line(aes(y = f, color = 'blue')) +
  geom_ribbon(
    aes(
      ymin = f - 1.96 * sqrt(Q),
      ymax = f + 1.96 * sqrt(Q)
    ),
    alpha = 0.2
  ) +
  scale_color_manual(
    labels = c("Predicted", "Observed"),
    values = c("red", "blue")
  ) +
  labs(
    title = expression(paste(
      "Forecasts of ",
      y[t],
      " w/ Standard Errors"
    )),
    y = "Observations and Forecasts",
    x = "Time",
    color = NULL
  )
```

We report the following $q_{101:110}$ values:

```{r echo=FALSE, results='asis'}
x[101:110,] %>%
  select(time, Q) %>%
  knitr::kable(align = 'll')
```

It can be seen above that the variance of the forecast increases with time. The reason for this is that the farther from the last data point that we forecast, the less confident we are in our prediction as we don't have a nearby reference to compare to.
