---
title: "Formulas and Functions"
output: html_document
date: '2022-06-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Conditional Distribution

Compute conditional mean:

```{r}
93 + t(c(0.1, 2)) %*% solve(matrix(c(9, 2, 2, 9), ncol = 2)) %*% c(2, -1)
```

Compute conditional variance:

```{r}
9 - t(c(0.1, 2)) %*% solve(matrix(c(9, 2, 2, 9), ncol = 2)) %*% c(.1, 2)
```

Compute area under bell curve:

```{r}
1-pnorm(95, 92.68831, sqrt(8.541688))
```

Generate white noise:

```{r}
w <- rnorm(500)
```

Moving average model (1/3 (wt-1 + w + wt+1)) on white noise:
We disambiguate from dplyr's filter function
```{r}
v <- stats::filter(w, sides = 2, filter = rep(1/3, 3))
```

Autoregression (with lag 2):
```{r}
x <- stats::filter(w, filter = c(1, -0.5), method = "recursive")[-(1:50)]
```

Random walk (autoregresssion with lag 1):
```{r}
rw <- cumsum(w)
```

Random walk with drift (autoregression with lag 1 and constant delta):
```{r}
wd <- w + .2
rwd <- cumsum(wd)
```

Signal plus noise:
In this example we add w to a standard wave function
```{r}

```

Signal-to-noise ratio (SNR):
ratio of amplitude over standard deviation of white noise
This example is A=2 and sigmaw=1 so SNR=2
```{r}

```

Gaussian Kernel Smoother
```{r}
k_smoothed_data <- data.frame(
  SOI = ksmooth(time(soi),
                as.numeric(soi),
                kernel = "normal",
                bandwidth = 1)$y,
  Time = as.Date(time(soi))
)
```

ARIMA Model
```{r}
x <- arima.sim(list(order = c(1, 1, 0), ar = 0.9), n = 150)[-1]
x_train <- window(x, start = 1, end = n_train)
fit_for <- sarima.for(x_train, n.ahead = 50, 1, 1, 0, plot = F)
```

Exponential Parametric Covariance Functions
```{r}
exp_cov_func <- function(i, j, sig, l) {
  sig*exp(-abs(i - j)*l)
}

sq_exp_cov_func <- function(i, j, sig, l) {
  sig*exp(-((abs(i - j)*l)^2))
}

sig_calc <- function(df, sigma2) {
  n = dim(df)[1]
  m = as.matrix(df)
  sigma2 - m[1,2:n] %*% solve(m[2:n,2:n]) %*% m[2:n,1]
}

mu_calc <- function(df, preds) {
  n = dim(df)[1]
  m = as.matrix(df)
  matrix(preds, nrow = 1, ncol = n-1) %*% solve(m[2:n,2:n]) %*% m[2:n,1]
}

build_cov_matrix <- function(data, func, sigma, l) {
  n = length(data)
  y_df <- data.frame(matrix(ncol = n, nrow = 0))
  for (i in 1:n) {
    new_row = c()
    for (j in 1:n) {
      new_row = c(new_row, func(data[i], data[j], sigma, l))
    }
    
    y_df[nrow(y_df) + 1,] = new_row
  }
  y_df
}

y <- c(0.6, 0.2, 0.3, 0.9)

y_df <- build_cov_matrix(y, sq_exp_cov_func, 10, 1)
mu_calc(y_df, c(3, 4, -1))
sqrt(sig_calc(y_df, 10))

y_df <- build_cov_matrix(y, sq_exp_cov_func, 10.01, 1)
mu_calc(y_df, c(3, 4, -1))
sqrt(sig_calc(y_df, 10.01))
```