---
title: "Formulas and Functions"
output: html_document
date: '2022-06-08'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Conditional Distribution

Compute conditional mean:

```{r}
93 + t(c(0.1, 2)) %*% solve(matrix(c(9, 2, 2, 9), ncol = 2)) %*% c(2, -1)
```

Compute conditional variance:

```{r}
9 - t(c(0.1, 2)) %*% solve(matrix(c(9, 2, 2, 9), ncol = 2)) %*% c(.1, 2)
```

Compute area under bell curve:

```{r}
1-pnorm(95, 92.68831, sqrt(8.541688))
```

Generate white noise:

```{r}
w <- rnorm(500)
```

Moving average model (1/3 (wt-1 + w + wt+1)) on white noise:
We disambiguate from dplyr's filter function
```{r}
v <- stats::filter(w, sides = 2, filter = rep(1/3, 3))
```

Autoregression (with lag 2):
```{r}
x <- stats::filter(w, filter = c(1, -0.5), method = "recursive")[-(1:50)]
```

Random walk (autoregresssion with lag 1):
```{r}
rw <- cumsum(w)
```

Random walk with drift (autoregression with lag 1 and constant delta):
```{r}
wd <- w + .2
rwd <- cumsum(wd)
```

Signal plus noise:
In this example we add w to a standard wave function
```{r}

```

Signal-to-noise ratio (SNR):
ratio of amplitude over standard deviation of white noise
This example is A=2 and sigmaw=1 so SNR=2
```{r}

```

Gaussian Kernel Smoother
```{r}
k_smoothed_data <- data.frame(
  SOI = ksmooth(time(soi),
                as.numeric(soi),
                kernel = "normal",
                bandwidth = 1)$y,
  Time = as.Date(time(soi))
)
```

ARIMA Model
```{r}
x <- arima.sim(list(order = c(1, 1, 0), ar = 0.9), n = 150)[-1]
x_train <- window(x, start = 1, end = n_train)
fit_for <- sarima.for(x_train, n.ahead = 50, 1, 1, 0, plot = F)
```