---
title: "Homework 1 Question 1"
author: "Rick Gray"
date: '2022-06-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
```

X_t = 10sin(t/5) + w_t

### Part A

We first solve for if mu is constant:

mu = E[X_t]\
mu = E[10sin(t/5) + w_t]\
mu = E[10sin(t/5)] + E[w_t]\
mu = E[10sin(t/5)] + 0\
mu = 10 * E[sin(t/5)]\
mu = 10 * 0\
mu = 0\

mu does not depend on time so this check is valid.

Next we solve for if the covariance depends only on time, or:\
gamma(s, t) = gamma(abs(s - t))

gamma_x(s, t) = cov(X_s, X_t)\
gamma_x(s, t) = cov(10sin(s/5) + w_s, 10sin(t/5) + w_t)\
gamma_x(s, t) = cov(10sin(s/5), 10sin(t/5)) + cov(w_s, w_t)\
gamma_x(s, t) = 100 * cov(sin(s/5), sin(t/5)) + sigma^2\
gamma_x(s, t) = 100 * E[(s - E[sin(s/5)])(t - E[sin(t/5)])] + sigma^2\
gamma_x(s, t) = 100 * E[(s - 0)(t - 0)] + sigma^2\
gamma_x(s, t) = 100 * E[st] + sigma^2\

This is not a function of abs(s - t), therefore it does not satisfy the second condition.

This signal plus noise process is NOT stationary.

### Part B

Signal to noise ratio = Amplitude of singal / variance of noise

```{r}
10 / 1
```

### Part C

```{r}
s <- 10*sin(1:120/5)
swd <- data.frame("Time" = 1:120, "x" = s)

ggplot(swd, aes(x = Time, y = x)) + 
  geom_line(size = 1, color = "blue") +
  ggtitle("Sine")

s <- 10*sin(1:120/5)

sw <- ts(s + rnorm(120))
swd <- data.frame("Time" = 1:120, "x" = sw)

low_noise <- ggplot(swd, aes(x = Time, y = x)) + 
  geom_line(size = 1, color = "blue") +
  geom_point(shape = 19, size = 2, fill = "white", color = "blue") +
  geom_hline(yintercept = 0,
             color = "darkgrey") +
  geom_hline(yintercept = mean(swd$x),
             linetype="dashed", 
             color = "blue") +
  scale_y_continuous(
    limits = c(-15, 15)
  ) +
  ggtitle("Signal with Low Noise")

low_noise
```

### Part D

```{r fig.width = 10}
sw2 <- ts(s + rnorm(120, 0, 16))
swd2 <- data.frame("Time" = 1:120, "x" = sw2)

high_noise <- ggplot(swd2, aes(x = Time, y = x)) + 
  geom_line(size = 1, color = "blue") +
  geom_point(shape = 19, size = 2, fill = "white", color = "blue") +
  geom_hline(yintercept = 0,
             color = "darkgrey") +
  geom_hline(yintercept = mean(swd$x),
             linetype="dashed", 
             color = "blue") +
  scale_y_continuous(
    limits = c(-15, 15)
  ) +
  ggtitle("Signal with High Noise")

plot_grid(low_noise, high_noise)
```

The new signal to noise ration is

```{r}
10 / 16
```

### Part E