---
title: "Homework 5 Question 1"
output: html_document
date: '2022-06-25'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(fig.width = 10)
library(tidyverse)
library(scales)
library(sf)
library(geoR)
```

### Part A

```{r}
rain <- read.table("rain.txt", header=T)
rain <- rain |>
  mutate(
    altitude = altitude / 5280,
    sqrt_rainfall = sqrt(rainfall)
  )

rain |>
  ggplot(aes(x, y, color = rainfall)) +
  geom_point(size = 2) +
  coord_fixed(ratio = 1) +
  scale_color_continuous(name = "Rainfall (mm)") +
  labs(
    title = "Average Yearly Rainfall in Switzerland Scatterplot",
    x = "X",
    y = "Y"
  )

rain |>
  ggplot(aes(x, y, color = altitude)) +
  geom_point(size = 2) +
  coord_fixed(ratio = 1) +
  scale_color_continuous(name = "Altitude (m)") +
  labs(
    title = "Altitude in Switzerland Scatterplot",
    x = "X",
    y = "Y"
  )
```

By visual inspection of the two plots, there appears to be some dependence on the amount of rainfall at a particular point, given its altitude. The top right of the plots are similar in darkness and seem to brighten downward and to the left.

### Part B

```{r}
lr <- lm(sqrt_rainfall~altitude, rain)
summary(lr)

```

The estimated regression equation is

<center>$y=14.127-8.368x$</center>\

and the estimated error variance is

<center>$var=4.275^2=18.276$</center>\

The altitude explains for the $\sqrt[2]{rainfall}$'s variation by a proportion of

<center>$R^2=0.03145$</center>

### Part C

```{r}
distances <- data.frame(matrix(ncol = 7, nrow = 0))
colnames(distances) <- c("x1", "y1", "resid1", "x2", "y2", "resid2", "distance")

for (i in 1:dim(rain)[1]) {
  for (j in 1:dim(rain)[1]) {
    if (i != j) {
      distances[nrow(distances) + 1,] = c(
        rain$x[i], rain$y[i], lr$residuals[i],
        rain$x[j], rain$y[j], lr$residuals[j],
        st_distance(st_as_sf(data.frame(
          x = c(rain$x[i], rain$x[j]),
          y = c(rain$y[i], rain$y[j])
        ), coords = c("x", "y")))[1, 2]
      )
    }
  }
}

distances |>
  ggplot(aes(distance, stat(count)/sum(stat(count)))) +
  geom_histogram(binwidth = 20, center = 10) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Frequency of Distances Between Coordinates",
    x = "Distance (m)",
    y = "Count"
  )
```

### Part D

```{r}
resid_corr <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(resid_corr) <- c("bin", "corr", "count")

for (b in seq(0, 280, 20)) {
  resid_corr[nrow(resid_corr) + 1,] = c(
    b + 10,
    (distances |>
    filter(distance >= b, distance < b + 20) |>
    select(resid1, resid2) |>
    cor())[1, 2],
    distances |>
      filter(distance >= b, distance < b + 20) |>
      count()
  )
}

resid_corr_plot <- resid_corr |>
  ggplot(aes(bin, corr, color = count)) +
  geom_point(size = 2.5) +
  labs(
    title = "Correlation of Residuals Binned by Distance",
    x = "Distance",
    y = "Correlation",
    color = "Count"
  )

resid_corr_plot
```

### Part E

There is a limited area with which our data is studied. Given this, distances that approach the width of the window are less likely to appear. The data is a _biased_ sample for this reason. We can't observe distances larger than the window. The result of this is that we expect to observe more smaller distances than larger ones.

### Part F

```{r}
rain_geo <- as.geodata(
  cbind(
    rain$sqrt_rainfall,
    rain$altitude,
    rain$x,
    rain$y
  ),
  data.col = 1,
  covar.col = 2,
  covar.names = "altitude",
  coords.col = 3:4
)

fit <- likfit(
  rain_geo,
  trend = ~altitude,
  cov.model = "exponential",
  ini.cov.pars = c(1, 1),
  fix.nugget=T,
  nugget=0)

fit
```

We report the estimated intercept as

<center>$\beta_0=11.5970$</center>

the estimated slope as

<center>$\beta_1=0.0288$</center>

and the estimated covariance parameters as

<center>$\sigma^2=20.9733$ and $\phi=42.4066$</center>

### Part G

```{r}
resid_corr_plot +
  geom_function(
    aes(x = 0),
    fun = function(x) {
      exp(x/-42.4066)
    },
    color = "red",
    size = 1.25
  ) +
  labs(title = "Correlation of Residuals Binned by Distance With Exponential Fit")
```

### Part H

```{r}
summary(fit)
```

We report the AIC for the non-spatial regression as

<center>$AIC=578.3$</center>

and the AIC for the spatial regression as

<center>$AIC=503.5$</center>

### Part I

The information that would be needed to make predictions in unmonitored locations is the __altitude__ at the desired prediction locations. Our model is fitted both on coordinates _and_ altitude; while we can simply pick coordinates we would like to predict rainfall at, we would need to acquire the altitude at the specified location.
