---
title: "Homework 2 Question 3"
author: "Rick Gray"
date: '2022-06-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Part A

The bandwidth of $a_2$ is much greater than $a_1$. The result is that a larger area of the original dataset is averaged into the computed data point and is therefore pulled closer to the mean. This smooths the moving average line.

### Part B

```{r fig.width=10}
data <- read_csv("TexasAccidents.csv", col_types = "Dnc")
data[,"MA"] <- stats::filter(
  data$Freq,
  sides = 2,
  filter = c(rep(c(1/7, rep(0, 6)), 6), 1/7)
)
data <- data[22:1072,]

data %>%
  ggplot(aes(Date, MA, color = DayOfWeek)) + 
  geom_line(size = .5) +
  geom_point(shape = 21, size = 1, fill = "black", color = "black") +
  labs(
    title = "Moving Average Filter of the Number of RecordedCar Crashes Each Day in Texas from 2017 to 2020",
    x = "Time",
    y = "Number of Recorded Car Crashes",
    color = "Day of Week"
  )
```

The plot shown above is significantly more effective at conveying the cycle that appears every other year and the effect of the day of the week. The weights in $a_3$ do not align well with the cycles in the data, while the weights in $a_4$ appear to fit the data quite well.

### Part C

The plot above is a good representation of why a simple visual inspection can be misleading. The frequency of car accidents follows a cycle that aligns with an event (perhaps an election), that occurs biannually on weekends. This would not be discernible without proper exploratory data analysis.