---
title: "Homework 4 Question 1"
output: html_document
date: '2022-06-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.width = 10)
library(tidyverse)
library(ggmap)
library(spatstat)
```

### Part A

```{r}
x <- read_csv("chi_markets.csv", show_col_types = F)

ggmap(get_map("Chicago")) +
  geom_point(aes(lon, lat), data = x) +
  labs(
    title = "Market Locations in City of Chicago",
    x = NULL,
    y = NULL
  )
```

There would not be any supermarkets on the top right corner of the map as this is Lake Michigan. It would not be possible to build a building on water.

### Part B

```{r}
load("chi_supermarkets_ppp.Rdata")
plot(markets_ppp, main="Market Locations and City Boundaries for Chicago")
```

While the data does appear to exhibit events with some regularity, there is a bit of clustering to the north east. The higher intensity in this area may lead us to believe that the data does arise from a homogeneous Poisson process. Because the data is not particularly regular, nor particularly clustered, it could be said to be Complete Spatial Randomness (CSR).

### Part C

```{r}
fit <- ppm(markets_ppp, ~ polynom(x, y, 2))

plot(markets_ppp, main="Market Locations and City Boundaries for Chicago")
plot(
  fit,
  add = TRUE,
  how = "image",
  se = FALSE,
  col = grey(seq(1, 0, length=128))
)

fit$coef[1]
predict(fit, locations = data.frame(x = 443674, y = 4636999))
```

We report the intercept of the estimated log intensity function to be

<center>$-31395.97$</center>\

We report the numerical value of the intensity function at $(443674,4636999)$ to be

<center>$7.7057{\times}10^{-7}$</center>